Finding a closest point in a lattice of Voronoi's first kind

### 2014-07-11 reviews/response.tex

``Motivated by Shalvi~et.~al.~\cite{Shalvi_iterativeslicer_2009} and  Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013}, we consider the following iteration,''

Among the iterative proceedures in our paper, in~\cite{Shalvi_iterativeslicer_2009}, and in~\cite{MicciancioVoulgaris_deterministic_jv_2013}, ours is the simplest.  The choice of our iteration is not motivated by what might be fastest in general lattices.  It is motivated by what we can prove to be polynomial time for those lattices of Vornoi's first kind.

The~\emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009} is similar but uses the (presumably accelerated?) iteration
\begin{align*}
x_{k+1} &= x_k + \alpha_k v_k  \\
(v_k, \alpha_k) &= \arg\min_{ (v, \alpha) \in ( \relevant(\Lambda) \cup \{\zerobf\}, \ints) } \|y - x_k - \alpha v \|.
\end{align*} 
So it allows adding integer multiples of a relevant vector.  We do not believe using this will improve the complexity of our algorithm beyond $O(n^4)$.  It would greatly complicate our proofs.

The iterative algorithm of Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} is different again.  To our (admittedly rough) knowledge, the proceedure computes an $\alpha_k$ (different, but related to that above) and then proceeds to compute a closest point in the scaled lattice $\alpha_k \Lambda$.  The scaling $\alpha_k$ is chosen so that the target point lies withing $2\alpha_k \vor(\Lambda)$.  Once a closest lattice point in $\alpha_k\Lambda$ is found, the next $\alpha_{k+1}$ is computed and the proceedure repeats.

We beleive that detailing the differences between these iterative proceedures would only detract from the core purpose of our paper to decscribe an $O(n^4)$ time algorithm to compute a closest point in lattice of Voronoi's first kind.

@@

\item\textbf{Comment:}  
page 5, as before, you should probably mention also Voulgaris when citing 37. 
\\\textbf{Response:}
Done.

\item\textbf{Comment:}  
page 6, Theorem 4.1 (this follows from [ 9, Theorem 2+3], not just theorem 3. 
Also, make the theorem statement self contained by adding "Let L be a lattice of voronoi first kind with obtuse superbasis b1..bn+1". 
\\\textbf{Response:}

\item\textbf{Comment:}  
page 8, lemma 5.1. I think you forgot a factor 2 preceding the double summations over S and T (at end of lemma statement, and also in the proof.) 
\\\textbf{Response:}
This error has been fixed.

\item\textbf{Comment:}  
page 11, ``with be'' $\to$ ``will be'' Displayed equation before ``Putting p=...'', A | should be || (5.7) again, missing factor 2 before double summation, and second set should be T, not S. 
\\\textbf{Response:}
These errors have been fixed.

@@

\item\textbf{Comment:}  
page 12, first displayed equation, usual problem: missing factor 2, and summation over S\&S instead of S\&T both on line 2 and 3. 
\\\textbf{Response:}
These errors have been fixed.

\item\textbf{Comment:}  
page 12, last line, definition of $s_i$, missing (-2) factor. 
\\\textbf{Response:}
This error has been fixed.

\item\textbf{Comment:}  
page 13, after (6.1), ``be computing'' $\to$ ``by computing'' 
\\\textbf{Response:}
Fixed.

\item\textbf{Comment:}  
page 14, ``have no effect'' $\to$ ``has no effect'' 
\\\textbf{Response:}
Fixed.

@@

\item\textbf{Comment:}  
page 15, definition of $Q$, should the summation defining $q_{ii}$ be restricted to $j\neq i$? 
\\\textbf{Response:}
Yes. This has been fixed.

\item\textbf{Comment:}  
page 15, ``there many'' $\to$ ``there may'' 
\\\textbf{Response:}
Fixed.

\item\textbf{Comment:}  
References: 
- 12: missing author (R. Raz) 
- 18: ``A V'' $\to$ ``A. V.'' etc, 
- 37: journal version in SIAM J. computing, 42(3):1364-1391 (2013) 
\\\textbf{Response:}
Fixed.

### 2014-07-11 reviews/response.tex

\item\textbf{Comment:}  
page 15, ``there many'' $\to$ ``there may'' 
\\\textbf{Response:}
Fixed.

\item\textbf{Comment:}  
References: 
- 12: missing author (R. Raz) 
- 18: ``A V'' $\to$ ``A. V.'' etc, 
- 37: journal version in SIAM J. computing, 42(3):1364-1391 (2013) 
\\\textbf{Response:}
Fixed.

\item\textbf{Comment:}  
Some additional references on CVPP you may find interesting 
(cite at your own discretion):

- O. Regev (2004) Improved Inapproximability of Lattice and Coding Problems With Preprocessing. IEEE Transactions on Information Theory 50(9): 2031-2037 
- ``Hardness of Approximating the Closest Vector PRoblem with Pre-Processing'', M. Alekhnovich, S. Khot, G. Kindler, N. Vishnoi, Computational Complexity 20(4):741-753, 2011.

@@

- D. Aharonov and Oded Regev (2005). Lattice problems in NP ∩ co-NP. J. ACM 52, 749-765.

(Gives polynomial time solution to approximate the distance estimation version of CVPP within $\gamma=\sqrt(n \log n)$ approximation factor, i.e., for any lattice there is an appropriate advice (more complex than just a good basis) that allows to approximate in polynomial time the distance of any target point to the lattice within gamma, but without actually finding a nearby lattice point.)

- On the Closest Vector Problem with a Distance Guarantee 
D. Dadush, O. Regev, N. Stephens-Davidowitz, Computational Complexity Conference 2014, to appear (next week!)

I am not sure, as the paper has not appeared yet, but I suspect it is the same as the paper on Dadush webpage with title ``On BDD and CVPP''. 
See link 
http://cs.nyu.edu/~dadush/papers/cvpp.pdf 
and read abstract for more information.
\\\textbf{Response:}
We thank the reviewer for these interesting references highlighting the difficultly of computing close approximations of a closest lattice point.  These have been referenced in Section~\zref{sec:iterative-slicer} with the following text:

### 2014-07-11 reviews/response.tex

The formula for $\|B\xbf\|_2^2$ stated by the reviewer is known as~\emph{Selling's formula} in the literature on quadratic forms and the geometry of numbers~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  We do indirectly use this formula in the form of the function $\Phi(S,p)$ and our Lemma~\zref{lem:decSellings}.  A very early draft of our paper attempted to use Selling's formula more directly, but we ultimately found that it was $\Phi(S,p)$ that was used repeatedly and not Selling's formula.

\item\textbf{Comment:} 
To present the result from the perspective of Computational Complexity (also not given in the
paper), this paper can be viewed as giving a polynomial time algorithm for the Closest Vector
Problem with Preprocessing (CVPP) over a special class of lattices. In the preprocessing model,
we are allowed to compute any polynomial amount of advice about the lattice (i.e. a good basis, or many short vectors in the lattice, etc.) before answering any CVP queries, where the resources needed to compute this advice are not counted in the runtime of the algorithm. In its
gap decisional version, where we need only decide whether a target is at distance $\leq d$ or $> \alpha d$
(for some approximation factor $\alpha \geq 1$), it is was shown [1, 4] to be NP-Hard for any constant $\alpha = 2^{\log^{1-\epsilon}n}$, and to not be polynomial time solvable for $\alpha = 2 \log n$ , for any fixed $\epsilon > 0$, under the assumption  that $NP \subsetneq \text{RTIME}(2^{\log^{O(1/\epsilon)}n})$ (quasi-polynomial time). For its approximate search version (i.e. find a lattice vector whose distance to the target is some bounded factor away from optimal), an $O(n^{1.5})$ approximation algorithm was (implicitly) given in [5], which was only recently improved to $O( \sqrt{ n / \log n} )$ in [3].
\\\textbf{Response:}

The following paragraph has been included in the introduction to put our results into context with the closest vector problem with preprocessing:

``Our results can be placed in the context of a modification of the closest lattice point problem called the \emph{closest vector problem with preprocessing}~\cite{micciancio_hardness_2001,feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aleknovish_hardness_with_preprocessing_2011,knot_hardness_cvvp_2014,Dadush_cvp_with_distance_guarantee_2014}.  In this problem some ``advice'' of polynomial size about the lattice is assumed to be given.  The advice might come in the form of a particular basis for the lattice or might take other forms.  The advice may be used to compute a closest lattice point, hopefully with reduced complexity.  %Even this modification of the closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001,feige_inapproximability_2004}.  
Our algorithm can be viewed as an efficient solution for the closest vector problem with preprocessing for the lattices of Voronoi's first kind.  The advice given is the obtuse superbasis.''

@@

The following paragraph has been included in the introduction to put our results into context with the closest vector problem with preprocessing:

``Our results can be placed in the context of a modification of the closest lattice point problem called the \emph{closest vector problem with preprocessing}~\cite{micciancio_hardness_2001,feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aleknovish_hardness_with_preprocessing_2011,knot_hardness_cvvp_2014,Dadush_cvp_with_distance_guarantee_2014}.  In this problem some ``advice'' of polynomial size about the lattice is assumed to be given.  The advice might come in the form of a particular basis for the lattice or might take other forms.  The advice may be used to compute a closest lattice point, hopefully with reduced complexity.  %Even this modification of the closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001,feige_inapproximability_2004}.  
Our algorithm can be viewed as an efficient solution for the closest vector problem with preprocessing for the lattices of Voronoi's first kind.  The advice given is the obtuse superbasis.''

\item\textbf{Comment:}  
Additional comments about the above paragraph: one author is missing in the reference for 12. Of [1,23,40] only 23 achieves $n^{O(n)}$ running time. (I think the algorithms in 40 take $2^{O(n^2)}$.) There is a journal version of 37 in SIAM J. Computing, and the paper is by two authors (not just Micciancio). 
\\\textbf{Response:}
We have fixed these errors.  This paragraph now reads:

``The closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.  These algorithms require a number of operations that grows as $O(n^{O(n)})$ or $O(n^{O(n^2)})$ where $n$ is the dimension of the lattice.  Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.  This single exponential growth in complexity is the best known.''

@@

``Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such the integer lattice $\ints^{n}$, the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, and the related Coxeter lattices~\cite[Chap.~4]{SPLAG}\cite{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices}.''

The interested reader could read these references in the order given. %, i.e., first Chapter 4 of~\cite{SPLAG} and~\cite{Conway1982FastQuantDec} for $D_n$ and $D_n^*$ and older algorithms for $A_n$ and $A_n^*$.  Then~\cite{McKilliam2008} for a fast algorithm for $A_n^*$, then~\cite{McKilliam2009CoxeterLattices} for fast algorithms for all of the Coxeter lattices that includes $A_n$ and $A_n^*$, $E_8$, etc.  Technically, the algorithm from~\cite{McKilliam2009CoxeterLattices} does include that from~\cite{McKilliam2008}.  However~\cite{McKilliam2008} is simpler and shorter to read first.

\item\textbf{Comment:}  
When saying ``if the obtuse superbasis is known'', you may want to cast this requirement in the context of CVPP. (The obtuse superbasis is the advice provided by the preprocessing.) 
\\\textbf{Response:}

\item\textbf{Comment:}  
page 2,bottom, displayed equation "$v \cdot x \leq x \cdot x$". Maybe you mean strict inequality.
\\\textbf{Response:}
%In our original definition we intended to include what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  This we not necessary
Yes, this has been fixed.  The text now reads reads:

The \emph{relevant vectors} are those lattice points $v \in \Lambda \backslash  \{\zerobf\}$ for which  
\[
\dotprod{v}{x} < \dotprod{x}{x} \qquad \text{for all $x \in \Lambda \backslash  \{\zerobf\}$}.
\]

% Our definition includes what Conway and Sloane call the `lax' relevant vectors.  To make this clear the following sentences have been added below the definition of the relevant vectors in Section~\zref{sec:voron-cells-relev}:

% ``This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.'' %It will be of benefit to include all of the strict and lax relevant vectors here.''

@@

\item\textbf{Comment:}  
page 5, as before, you should probably mention also Voulgaris when citing 37. 
\\\textbf{Response:}
Done.

\item\textbf{Comment:}  
page 5, last line, "dimension less than 4", does it include 4 or not? 
\\\textbf{Response:}
It does not include 4.  We have modified the text here to read ``less than or equal to $3$''. We have also modified similar text in the last paragraph of Section~\zref{sec:discussion} to read ``It is known that all lattices of dimension less than or equal to 3 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.  Lattices of dimension larger than 3 need not be of Voronoi's first kind.''

\item\textbf{Comment:}  
page 5, as before, you should probably mention also Voulgaris when citing 37. 
\\\textbf{Response:}
Done.

\item\textbf{Comment:}  
page 6, Theorem 4.1 (this follows from [ 9, Theorem 2+3], not just theorem 3. 
Also, make the theorem statement self contained by adding "Let $L$ be a lattice of voronoi first kind with obtuse superbasis $b_1,\dots,b_{n+1}$". 
\\\textbf{Response:}
We beleive that it is only Theorem 3 from~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} that needs to cited.  Theorem~2 in \cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} asserts that the relevant vectors in any lattice $\Lambda$ are the shortest coset representatives of the quotient $\Lambda/2\Lambda$.  Theorem~3 uses this, Selling's formula, and the fact that the Selling parameters are nonpositive to prove that the relevant vectors in a lattice of Voronoi's first kind take the form
\[
\sum_{i \in I} b_i
\]
where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty.

We have included the text ``Let $\Lambda$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.'' into the statement of Theorem 4.1.

%BLERG: Check that Theorems 2 and 3 in the published version of\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} are what I think they are!

### 2014-07-11 paper.tex
minor

The closest lattice point problem and the shortest vector problem have interested mathematicians and computer scientists due to their relationship with integer programming~\cite{Lenstra_integerprogramming1983,Kannan1987_fast_general_np,Babai1986}, the factoring of polynomials~\cite{Lenstra1982}, and cryptanalysis~\cite{Joux_toolbox_cryptanal1998,NyguyenStern_two_faces_crypto,Micciancio_lattice_based_post_quantum_crypto}.  
Solutions of the closest lattice point problem have engineering applications.  For example, if a lattice is used as a vector quantiser then the closest lattice point corresponds to the minimum distortion point~\cite{Conway1983VoronoiCodes,Conway1982VoronoiRegions,Conway1982FastQuantDec}.  If the lattice is used as a code, then the closest lattice point corresponds to what is called \emph{lattice decoding}\index{lattice decoding} and has been shown to yield arbitrarily good codes~\cite{Erex2004_lattice_decoding,Erez2005}.  The closest lattice point problem also occurs in communications systems involving multiple antennas~\cite{Ryan2008,Wubben_2011}.  The unwrapping of phase data for location estimation can also be posed as a closest lattice point problem and this has been applied to the global positioning system~\cite{Teunissen_GPS_1995,Hassibi_GPS_1998}.  The problem has also found applications to circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}, single frequency estimation~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, and related signal processing problems~\cite{McKilliam2007,Clarkson2007,McKilliam2009IndentifiabliltyAliasingPolyphase,Quinn_sparse_noisy_SSP_2012}.

The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.  These algorithms require a number of operations that grows as $O(n^{O(n)})$ or $O(n^{O(n^2)})$ where $n$ is the dimension of the lattice.  Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.  This single exponential growth in complexity is the best known.

@@

@@

Equivalently, the Voronoi cell can be defined as the intersection of the half spaces 
\begin{align*}
H_{v} &= \{x \in \reals^n \mid \|x\| \leq \|x - v\| \} \\
&= \{x \in \reals^n \mid \dotprod{x}{v} \leq \tfrac{1}{2}\dotprod{v}{v} \}
\end{align*}
for all $v \in \Lambda \backslash  \{\zerobf\}$.  %We denote by $x^\prime$ the transpose of the vector $x$ and so $\dotprod{x}{v}$ is the inner product between column vectors $x$ and $v$.  
We denote by $\dotprod{x}{v}$ the inner product between vectors $x$ and $v$.
It is not necessary to consider all $v \in \Lambda \backslash  \{\zerobf\}$ to define the Voronoi cell.   %Figure \ref{lattices:fig:vorregion} depicts the relevant vectors of the lattice with generator given by~\eqref{lattices:eq:genB}.  A lattice point in $\calR$ is called \term{relevant}. The following remark follow trivially from the preceding discussion.
The \emph{relevant vectors} are those lattice points $v \in \Lambda \backslash  \{\zerobf\}$ for which  
\[
\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$}.
\]
This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered. %It will be of benefit to include all of the strict and lax relevant vectors here.  
Denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
\[
%\begin{equation}\label{eq:defvortermsrelvecs}
\vor(\Lambda) = \cap_{v\in\relevant(\Lambda)}{H_{v}}.
%\end{equation}
\]
The closest lattice point problem and the Voronoi cell are related in that $x\in\Lambda$ is a closest lattice point to $y$ if and only $y - x \in \vor(\Lambda)$, that is, if and only if
\begin{equation}\label{eq:relvectnearpointieq}
\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v} \qquad \text{for all $v \in \relevant(\Lambda)$}.  
\end{equation}

@@

\begin{theorem} \label{thm:revvecssuperbase} (Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases})
The relevant vectors of $\Lambda$ are of the form,
\[
\sum_{i \in I} b_i
\]
where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty, i.e., $I \subset \{1, 2, \dots, n+1\}$ and $I \neq \emptyset$.
\end{theorem}

>>>>

The closest lattice point problem and the shortest vector problem have interested mathematicians and computer scientists due to their relationship with integer programming~\cite{Lenstra_integerprogramming1983,Kannan1987_fast_general_np,Babai1986}, the factoring of polynomials~\cite{Lenstra1982}, and cryptanalysis~\cite{Joux_toolbox_cryptanal1998,NyguyenStern_two_faces_crypto,Micciancio_lattice_based_post_quantum_crypto}.  
Solutions of the closest lattice point problem have engineering applications.  For example, if a lattice is used as a vector quantiser then the closest lattice point corresponds to the minimum distortion point~\cite{Conway1983VoronoiCodes,Conway1982VoronoiRegions,Conway1982FastQuantDec}.  If the lattice is used as a code, then the closest lattice point corresponds to what is called \emph{lattice decoding}\index{lattice decoding} and has been shown to yield arbitrarily good codes~\cite{Erex2004_lattice_decoding,Erez2005}.  The closest lattice point problem occurs in communications systems involving multiple antennas~\cite{Ryan2008,Wubben_2011}.  The unwrapping of phase data for location estimation can be posed as a closest lattice point problem and this has been applied to the global positioning system~\cite{Teunissen_GPS_1995,Hassibi_GPS_1998}.  The problem has also found applications to circular statistics~\cite{McKilliam_mean_dir_est_sq_arc_length2010}, single frequency estimation~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, and related signal processing problems~\cite{McKilliam2007,Clarkson2007,McKilliam2009IndentifiabliltyAliasingPolyphase,Quinn_sparse_noisy_SSP_2012}.

The closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}.  Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.  These algorithms require a number of operations that grows as $O(n^{O(n)})$ or $O(n^{O(n^2)})$ where $n$ is the dimension of the lattice.  Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.  This single exponential growth in complexity is the best known.

@@

Our results can be placed in the context of a modification of the closest lattice point problem called the \emph{closest vector problem with preprocessing}~\cite{micciancio_hardness_2001,feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aleknovish_hardness_with_preprocessing_2011,knot_hardness_cvvp_2014,Dadush_cvp_with_distance_guarantee_2014}.  In this problem some ``advice'' of polynomial size about the lattice is assumed to be given.  The advice might come in the form of a particular basis for the lattice or might take other forms.  The advice may be used to compute a closest lattice point, hopefully with reduced complexity.  %Even this modification of the closest lattice point problem is known to be NP-hard~\cite{micciancio_hardness_2001,feige_inapproximability_2004}.  
Our algorithm can be viewed as an efficient solution for the closest vector problem with preprocessing for the lattices of Voronoi's first kind.  The advice given is the obtuse superbasis.

@@

Equivalently, the Voronoi cell can be defined as the intersection of the half spaces 
\begin{align*}
H_{v} &= \{x \in \reals^n \mid \|x\| \leq \|x - v\| \} \\
&= \{x \in \reals^n \mid \dotprod{x}{v} \leq \tfrac{1}{2}\dotprod{v}{v} \}
\end{align*}
for all $v \in \Lambda \backslash  \{\zerobf\}$.  %We denote by $x^\prime$ the transpose of the vector $x$ and so $\dotprod{x}{v}$ is the inner product between column vectors $x$ and $v$.  
We denote by $\dotprod{x}{v}$ the inner product between vectors $x$ and $v$.
It is not necessary to consider all $v \in \Lambda \backslash  \{\zerobf\}$ to define the Voronoi cell.   %Figure \ref{lattices:fig:vorregion} depicts the relevant vectors of the lattice with generator given by~\eqref{lattices:eq:genB}.  A lattice point in $\calR$ is called \term{relevant}. The following remark follow trivially from the preceding discussion.
The \emph{relevant vectors} are those lattice points $v \in \Lambda \backslash  \{\zerobf\}$ for which  
\[
\dotprod{v}{x} < \dotprod{x}{x} \qquad \text{for all $x \in \Lambda \backslash  \{\zerobf\}$}.
\]
%This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered. %It will be of benefit to include all of the strict and lax relevant vectors here.  
Denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
\[
%\begin{equation}\label{eq:defvortermsrelvecs}
\vor(\Lambda) = \cap_{v\in\relevant(\Lambda)}{H_{v}}.
%\end{equation}
\]
The closest lattice point problem and the Voronoi cell are related in that $x\in\Lambda$ is a closest lattice point to $y$ if and only $y - x \in \vor(\Lambda)$, that is, if and only if
\begin{equation}\label{eq:relvectnearpointieq}
\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v} \qquad \text{for all $v \in \relevant(\Lambda)$}.  
\end{equation}

@@

\begin{theorem} \label{thm:revvecssuperbase} (Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}) Let $\Lambda$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.  The relevant vectors of $\Lambda$ are of the form,
\[
\sum_{i \in I} b_i
\]
where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty, i.e., $I \subset \{1, 2, \dots, n+1\}$ and $I \neq \emptyset$.
\end{theorem}
